---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r echo=FALSE}
setwd("~/a/highEd/dataScience_coursera/reproduce/prjct1")
   repTime<-format(Sys.time(), "%a %b %d %X %Y" )
```
*Running this report at `r repTime`*
```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

## Loading and preprocessing the data
```{r}
stepdata=read.csv("data/activity.csv")
summary(stepdata)
# Create a dataframe cmplt with no missing values
cmplt<-stepdata[!is.na(stepdata$steps),]
```
## Find the mean and median of the total number of steps per day
Since this will be done 2 times - once ignoring missing vlaues and
once with imputed missing values - the code to summarise the data
will be put in a reuseable function.
```{r}
stepsPerDay<-function(df) {
   steps=aggregate(df$steps, by=list(df$date), FUN=sum)
   
   step.mean=mean(steps$x)
   step.median=median(steps$x)
   tmp=sprintf("Steps per day.  Mean: %8.1f  Median %8.1f",
               step.mean, step.median)

   library(ggplot2)
   plt=ggplot(steps, aes(x=x)) + 
      geom_histogram(,color="black", fill="white", binwidth=2000) +
      labs(title=tmp, y="Number of days", 
           x="Number of Steps") +
      geom_vline(aes(xintercept=step.mean), color="red", 
                 linetype="dashed", size=1)
   plt
}
```
Create a histogram to report on the Steps per day when missing values
are not included.  Also report the Mean and Median steps per day.
```{r}
stepsPerDay(cmplt)
```
  

## What is the average daily activity pattern?

```{r}
avg5=aggregate(cmplt$steps, by=list(cmplt$interval), FUN=mean)
most4=avg5[avg5$x==max(avg5$x),]
tmp<-sprintf("%s\nMaximum steps (%4.1f) occured interval %5i", 
      "Average steps per day for each 5 minute interval",
      most4$x, most4$Group.1)

ggplot(avg5, aes(x=Group.1, y=x)) + geom_line() + 
   labs(title=tmp, x="(5 minute) Interval", y="Number of Steps")

```

## Imputing missing values
Make a dataframe like the original data but replace missing values
for steps with imputed values.  The imputed values are the average
number of steps for each time interval.
```{r}
nmiss=sum(is.na(stepdata$steps))
# create df impute like stepdata plus column "x" mean # steps @ interval
impute<-merge(stepdata, avg5, by.y="Group.1", by.x="interval")
# replace each missing steps with average sterps for that interval
impute$steps[is.na(impute$steps)]<-impute$x[is.na(impute$steps)]
summary(impute)
```
   
There were `r nmiss` missing values in the steps that have now been
replaced with the average values for that time period.  Re-run the 
caluculations of mean, median and the regnersate the histogram as was
done in the first step in this document.
  
```{r}
stepsPerDay(impute)
```
  
The effect of the impute is almost imperceptable, the median now equlas
the mean, but they were nearly equal to start with.

  

## Are there differences in activity patterns between weekdays and weekends?
